---
title: "Project4_6723_botelho3"
author: "botelho3"
date: '2022-12-04'
output:
  html_document:
    toc: yes
    df_print: paged
  html_notebook:
    theme: readable
    toc: yes
    toc_float: yes
---


```{r warning=FALSE, include=FALSE}
library(dplyr)
library(lubridate)
library(ggplot2)
library(recommenderlab)
library(DT)
library(data.table)
library(reshape2)
```

## Read Data


**Load Ratings Data**
```{r echo=FALSE}
# use colClasses = 'NULL' to skip columns
ratings = read.csv('ratings.dat', 
                   sep = ':',
                   colClasses = c('integer', 'NULL'), 
                   header = FALSE)
colnames(ratings) = c('UserID', 'MovieID', 'Rating', 'Timestamp')
head(ratings)
```

**Load Movies Data**
```{r echo=FALSE}
movies = readLines('movies.dat')
movies = strsplit(movies, split = "::", fixed = TRUE, useBytes = TRUE)
movies = matrix(unlist(movies), ncol = 3, byrow = TRUE)
movies = data.frame(movies, stringsAsFactors = FALSE)
colnames(movies) = c('MovieID', 'Title', 'Genres')
movies$MovieID = as.integer(movies$MovieID)

# convert accented characters
movies$Title = iconv(movies$Title, "latin1", "UTF-8")

# extract year
movies$Year = as.numeric(unlist(
  lapply(movies$Title, function(x) substr(x, nchar(x)-4, nchar(x)-1))))
head(movies)
```

**Load User Data**
```{r echo=FALSE}
users = read.csv('users.dat',
                 sep = ':', header = FALSE)
users = users[, -c(2,4,6,8)] # skip columns
colnames(users) = c('UserID', 'Gender', 'Age', 'Occupation', 'Zip-code')
head(users)
```


<!-- ## Exploration -->
<!-- ```{r} -->
<!-- dim(users) -->
<!-- length(unique(ratings$UserID)) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- movies_not_rated = movies %>%  -->
<!--   filter(!(MovieID %in% ratings$MovieID)) -->
<!-- dim(movies_not_rated) -->
<!-- ``` -->

<!-- **Distribution of Ratings** -->

<!-- ```{r} -->
<!-- tmp = data.frame(Rating = 1:5,  -->
<!--                  freq = as.vector(table(ratings$Rating)/nrow(ratings))) -->
<!-- ggplot(data = tmp, aes(x = Rating, y = freq)) + -->
<!--   geom_bar(stat="identity", fill = 'steelblue', width = 0.6) +  -->
<!--   geom_text(aes(label=round(freq, dig=2)),  -->
<!--                 vjust=1.6, color="white", size=3.5) + -->
<!--   theme_minimal() -->
<!-- ``` -->


<!-- **Ratings Per User** -->
<!-- ```{r} -->
<!-- tmp = ratings %>%  -->
<!--   group_by(UserID) %>%  -->
<!--   summarize(ratings_per_user = n())  -->
<!--   summary(tmp$ratings_per_user) -->
<!--   stem(tmp$ratings_per_user) -->
<!--   sum(tmp$ratings_per_user > 500) -->
<!--   sort(tmp$ratings_per_user[tmp$ratings_per_user>1300]) -->

<!-- tmp %>% -->
<!--   ggplot(aes(ratings_per_user)) + -->
<!--   geom_bar(fill = "steelblue") + coord_cartesian(c(20, 500)) -->

<!-- #tmp = tmp %>% full_join(users, by = 'UserID') -->
<!-- ``` -->

<!-- **Ratings Per Movie** -->

<!-- ```{r} -->
<!-- tmp = ratings %>%  -->
<!--   group_by(MovieID) %>%  -->
<!--   summarize(ratings_per_movie = n(), ave_ratings = mean(Rating)) %>% -->
<!--   inner_join(movies, by = 'MovieID') -->
<!-- summary(tmp$ratings_per_movie) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- tmp %>%  -->
<!--   filter(ratings_per_movie > 2000) %>% -->
<!--   arrange(desc = ratings_per_movie) %>% -->
<!--   select(c("Title", "ratings_per_movie")) %>% -->
<!--   print(n = 31) -->

<!-- tmp %>% ggplot(aes(ratings_per_movie)) +  -->
<!--   geom_bar(fill = "steelblue", width = 1) + coord_cartesian(c(1,1500)) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- small_image_url = "https://liangfgithub.github.io/MovieImages/" -->
<!-- ratings %>%  -->
<!--   group_by(MovieID) %>%  -->
<!--   summarize(ratings_per_movie = n(),  -->
<!--             ave_ratings = round(mean(Rating), dig=3)) %>% -->
<!--   inner_join(movies, by = 'MovieID') %>% -->
<!--   filter(ratings_per_movie > 1000) %>% -->
<!--   top_n(10, ave_ratings) %>% -->
<!--   mutate(Image = paste0('<img src="',  -->
<!--                         small_image_url,  -->
<!--                         MovieID,  -->
<!--                         '.jpg?raw=true"></img>')) %>% -->
<!--   select('Image', 'Title', 'ave_ratings') %>% -->
<!--   arrange(desc(-ave_ratings)) %>% -->
<!--   datatable(class = "nowrap hover row-border",  -->
<!--             escape = FALSE,  -->
<!--             options = list(dom = 't', -->
<!--                           scrollX = TRUE, autoWidth = TRUE)) -->
<!-- ``` -->


**Dist of Genres**

Compute matrix of Genres by MovieIDs.

```{r}
genres = as.data.frame(movies$Genres, stringsAsFactors=FALSE)
tmp = as.data.frame(tstrsplit(genres[,1], '[|]',
                              type.convert=TRUE),
                    stringsAsFactors=FALSE)
genre_list = c("Action", "Adventure", "Animation", 
               "Children's", "Comedy", "Crime",
               "Documentary", "Drama", "Fantasy",
               "Film-Noir", "Horror", "Musical", 
               "Mystery", "Romance", "Sci-Fi", 
               "Thriller", "War", "Western")
m = length(genre_list)
genre_matrix = matrix(0, nrow(movies), length(genre_list))
for(i in 1:nrow(tmp)){
  genre_matrix[i,genre_list %in% tmp[i,]]=1
}
colnames(genre_matrix) = genre_list
remove("tmp", "genres")
```

```{r}
data.frame(Genres = genre_list, 
                 Freq = as.vector(colMeans(genre_matrix))) %>% 
  ggplot(aes(reorder(Genres, Freq), Freq, fill = Freq)) + 
  geom_bar(stat = "identity") + 
    geom_text(aes(label = round(Freq, dig=2)), 
            position = position_stack(vjust = 0.5), 
            color="white", size=3) + 
  coord_flip() + 
  scale_colour_brewer(palette="Set1") + 
  labs(y = 'Frequency', x = 'Genre')
```


```{r}
tmp = ratings %>% 
  left_join(data.frame(MovieID = movies$MovieID, genre_matrix), 
            by = "MovieID") %>%
  select(-c("UserID", "MovieID", "Rating", "Timestamp"))
data.frame(Genres = genre_list, 
           Popularity = as.vector(colMeans(tmp))) %>% 
  ggplot(aes(reorder(Genres, Popularity), Popularity, fill = Popularity)) + 
  geom_bar(stat = "identity") + 
  geom_text(aes(label = round(Popularity, dig=3)), 
            position = position_stack(vjust = 0.5), 
            color="white", size=3) + 
  coord_flip() + 
  labs(y = 'Popularity', x = 'Genre')
```

```{r}
tmp = rowSums(genre_matrix)
summary(tmp)

movies[which(tmp==6), ]
movies[which(tmp==5), ]
```


```{r}
# range(movies$Year) % 1919 to 2000
tmp = data.frame(Year = movies$Year, genre_matrix) %>%
  group_by(Year) %>%
  summarise_all(sum)
tmp[,-1] = apply(tmp[, -1], 2, cumsum)
tmp[,-1] = tmp[,-1]/sum(tmp[nrow(tmp), -1])
print(round(tmp[nrow(tmp),-1], dig=3))
```

```{r}
tmp = reshape2::melt(tmp, id.vars="Year") 
tmp %>%
  ggplot(aes(Year, value, group = variable)) +
  geom_area(aes(fill = variable)) + 
  geom_line(aes(group = variable), position = "stack")
```






# System 1 - Recommendation Based on Genres


## Data Precprocessing

**Movies Data**
```{r}
movies = readLines('movies.dat')
movies = strsplit(movies, split = "::", fixed = TRUE, useBytes = TRUE)
movies = matrix(unlist(movies), ncol = 3, byrow = TRUE)
movies = data.frame(movies, stringsAsFactors = FALSE)
colnames(movies) = c('MovieID', 'Title', 'Genres')
movies$MovieID = as.integer(movies$MovieID)

# convert accented characters
movies$Title[73]
movies$Title = iconv(movies$Title, "latin1", "UTF-8")
movies$Title[73]

# extract year
movies$Year = as.numeric(unlist(
  lapply(movies$Title, function(x) substr(x, nchar(x)-4, nchar(x)-1))))
```

**User Data**
```{r}
users = read.csv('users.dat',
                 sep = ':', header = FALSE)
users = users[, -c(2,4,6,8)] # skip columns
colnames(users) = c('UserID', 'Gender', 'Age', 'Occupation', 'Zip-code')
```


## Recommendation Algorithm 1

Top five most highly rated movies in a genre.

```{r}
checkbox_genres = c(genre_list[1], genre_list[8], genre_list[length(genre_list)])
print(checkbox_genres)
system_1_recommend <- function(ratings, movies,
                     genre_list, genre_matrix,
                     checkbox_genres) {
  num_genres = length(checkbox_genres)
  TOP_N = 6
  joined = ratings %>% 
    group_by(MovieID) %>% 
    summarize(ratings_per_movie = n(), ave_ratings = mean(Rating))
  joined = movies %>%
    left_join(joined, by = 'MovieID', all.x = TRUE)# %>%
    #arrange(MovieID)
 
  # Transpose the relevant genres. 
  row_idx = which(genre_list %in% checkbox_genres) 
  genre_names = genre_list[row_idx]
  tgm = t(genre_matrix)
  tgm = tgm[row_idx, ]

  result_cols =  c("MovieID", "ave_ratings", "Genres", "Genre") 
  results_list = vector("list", num_genres) 
  names(results_list) = checkbox_genres
  results = data.frame(matrix(nrow = num_genres, ncol = 4))
  colnames(results) <- result_cols
  for (i in 1:length(row_idx)) {
    movies_row_idx = which(tgm[i, ] == 1)
    tmp = joined %>%
            slice(movies_row_idx) %>%
            mutate(mean_ratings_per_movie = mean(ratings_per_movie, na.rm = TRUE)) %>%
            filter(ratings_per_movie > mean_ratings_per_movie) %>%
            arrange(desc(ave_ratings)) 
    s = (i-1) * TOP_N + 1
    e = s + TOP_N - 1
    c = result_cols[!(result_cols %in% c("Genre"))]
    results[s:e, c] = tmp[1:TOP_N, c]
    results[s:e, ]$Genre = genre_names[i]
    results_list[[i]] = tmp$MovieID[1:TOP_N]
  }
  return(results)
}
# system_1_results = results
system_1_results = system_1_recommend(ratings, movies,
                                      genre_list, genre_matrix, checkbox_genres)

# Display results
for (g in checkbox_genres) {
  print(system_1_results[system_1_results$Genre == g, ])
}
```


## Recommendation Algorithm 2

Most popular movies in the genre.

- Highest number of average ratings. 
- Discount factor for age. Older movies are likely to have more ratings? So
discount them.

```{r}
# Add a ratings_per_movie and ave_ratings column
tmp = ratings %>% 
  group_by(MovieID) %>% 
  summarize(ratings_per_movie = n(), ave_ratings = mean(Rating)) %>%
  inner_join(movies, by = 'MovieID')
dim(tmp)

# Add a ratings_per_movie_year column
tmp = ratings %>% 
  inner_join(movies, by = 'MovieID') %>%
  group_by(Year) %>% 
  mutate(ratings_per_movie_year = n()) %>%
  group_by(MovieID) %>%
  summarize(ratings_per_movie = n(), ave_ratings = mean(Rating),
            ratings_per_movie_year = first(ratings_per_movie_year), Year = first(Year),
            Title = first(Title))
dim(tmp)

tmp %>%
  filter(ratings_per_movie > 2000) %>%
  arrange(desc(ratings_per_movie)) %>%
  select(c("Title", "ratings_per_movie", "ratings_per_movie_year", "Year")) %>%
  print(n = 31)

# unique((tmp %>% group_by(ratings_per_movie))$ratings_per_movie)
# unique((tmp %>% group_by(ratings_per_movie_year))$ratings_per_movie_year)

# Plot histogram of ratings_per_movie
tmp %>%
  group_by(ratings_per_movie) %>%
  mutate(count = n()) %>%
  ggplot(., aes(x = ratings_per_movie, y = count)) + 
  geom_bar(stat='identity', fill = "steelblue", width = 1) + coord_cartesian(c(20, 500))

#  # Same as above just different plotting method.
# tmp %>%
#   group_by(ratings_per_movie) %>%
#   ggplot(., aes(ratings_per_movie)) + 
#   geom_bar(fill = "steelblue", width = 1) + coord_cartesian(c(20, 500))

# Plot hisogram of movies per year
tmp %>%
  group_by(Year) %>%
  ggplot(aes(Year)) + 
    geom_bar(fill = "steelblue", width = 1) + coord_cartesian(c(1900,2020))


# Plot histogram of reviews_per_movie_year
tmp %>%
  group_by(Year) %>%
  summarize(ratings_per_movie_year = first(ratings_per_movie_year)) %>%
  ggplot(aes(x = Year, y = ratings_per_movie_year)) + 
    geom_bar(stat='identity', fill = "steelblue", width = 1) + coord_cartesian(c(1900,2020))
```

```{r}
# Add a ratings_per_movie and ave_ratings column
tmp = ratings %>% 
  group_by(MovieID) %>% 
  summarize(ratings_per_movie = n(), ave_ratings = mean(Rating)) %>%
  inner_join(movies, by = 'MovieID')
dim(tmp)

# Add a ratings_per_year column
tmp = ratings %>% 
  inner_join(movies, by = 'MovieID') %>%
  group_by(Year) %>% 
  mutate(ratings_per_movie_year = n()) %>%
  ungroup %>%
  group_by(MovieID) %>%
  summarize(ratings_per_movie = n(), ave_ratings = mean(Rating),
            ratings_per_movie_year = first(ratings_per_movie_year), Year = first(Year),
            Title = first(Title))

# tmp = tmp %>%
#   mutate(pct_ratings_per_year = (ratings_per_movie / ratings_per_movie_year)) %>%
#   arrange(desc(pct_ratings_per_year))

# Add a ratings_per_last_5_yr
last_year = 2003
ratings = ratings %>% 
  mutate(RYear = year(as_datetime(Timestamp)))
  # group_by(RYear) %>%
  # mutate(ratings_per_year = n())
head(ratings)
head(tmp)
```


```{r}
length(which(ratings$RYear == 2000))
length(which(ratings$RYear == 2001))
length(which(ratings$RYear == 2002))
length(which(ratings$RYear == 2003))

# prev = filter(RYear < last_year) %>%
#   inner_join(tmp, by = 'MovieID', suffix=c("", ".y")) %>%
#   select(-ends_with(".y")) %>%
#   group_by(Year) %>% 
#   mutate(mean_ratings_per_movie_year = n()) %>%
  
prev = ratings %>%
  filter(RYear < 2001) %>%
  inner_join(movies, by = 'MovieID') %>%
  group_by(MovieID) %>%
  mutate(ratings_per_movie = n(),
         ave_ratings = mean(Rating)) %>%
  ungroup %>%
  group_by(Year) %>% 
  mutate(ratings_per_movie_year = n()) %>%
  ungroup()
prev = prev %>%
  mutate(pct_ratings_per_movie_year =  ratings_per_movie / ratings_per_movie_year)

now = ratings %>%
  filter(RYear >= 2001) %>%
  inner_join(movies, by = 'MovieID') %>%
  group_by(MovieID) %>%
  mutate(ratings_per_movie = n(),
         ave_ratings = mean(Rating)) %>%
  ungroup %>%
  group_by(Year) %>% 
  mutate(ratings_per_movie_year = n()) %>%
  ungroup()
now = now %>%
  mutate(pct_ratings_per_movie_year =  ratings_per_movie / ratings_per_movie_year)
 

now = now %>% 
  group_by(MovieID) %>%
  summarize(
    MovieID = first(MovieID),
    Rating = mean(Rating),
    ratings_per_movie = first(ratings_per_movie),
    ratings_per_movie_year = first(ratings_per_movie_year),
    pct_ratings_per_movie_year = first(pct_ratings_per_movie_year),
    ave_ratings = first(ave_ratings),
  )
  
prev = prev %>% 
  group_by(MovieID) %>%
  summarize(
    MovieID = first(MovieID),
    Title = first(Title),
    Rating = mean(Rating),
    Year = first(Year),
    RYear = first(RYear),
    ratings_per_movie = first(ratings_per_movie),
    ratings_per_movie_year = first(ratings_per_movie_year),
    pct_ratings_per_movie_year = first(pct_ratings_per_movie_year),
    ave_ratings = first(ave_ratings),
  )
  
a =  prev %>%
  full_join(now, by = 'MovieID')

a = a %>%
  mutate(ratio = pct_ratings_per_movie_year.y / pct_ratings_per_movie_year.x) %>%
  mutate(diff = pct_ratings_per_movie_year.y - pct_ratings_per_movie_year.x) %>%
  mutate(ratio2 = ratings_per_movie.y / ratings_per_movie.x) %>%
  mutate(diff2 = ratings_per_movie.y - ratings_per_movie.x) %>%
  mutate(metric1 = scale(ratings_per_movie.y) - scale(ratings_per_movie.x))

to_order = a %>%
  select(MovieID, Title, ratio, diff, ratio2, diff2, metric1, Year, Rating.x) %>%
  filter(Rating.x > 4.0) %>%
  arrange(desc(metric1))
print(to_order, n=30)

head(tmp, n = 30)
head(prev[!duplicated(prev$MovieID), ], n = 30)
```
















# System 2 - Collaborative Recommendation System



```{r}
set.seed(6723)
library(recommenderlab)
library(proxy)
library(testit)
ratings = read.csv('ratings.dat', 
                   sep = ':',
                   colClasses = c('integer', 'NULL'), 
                   header = FALSE)
colnames(ratings) = c('UserID', 'MovieID', 'Rating', 'Timestamp')
i = paste0('u', ratings$UserID)
j = paste0('m', ratings$MovieID)
x = ratings$Rating
tmp = data.frame(i, j, x, stringsAsFactors = T)
Rmat = sparseMatrix(as.integer(tmp$i), as.integer(tmp$j), x = tmp$x)
rownames(Rmat) = levels(tmp$i)
colnames(Rmat) = levels(tmp$j)
M = dim(Rmat)[1]  # users
N = dim(Rmat)[2]  # ratings
Rmat = new('realRatingMatrix', data = Rmat)

train = Rmat[1:500, ]
test = Rmat[501, ]
```




### UBCF Recommendation

**My Implementation**

```{r}
set.seed(6723)
nn = 20
M = 500
N = N

# center
newdata = as(test, "matrix")
newuser.means = rowMeans(newdata, na.rm = TRUE)
newdata = newdata - newuser.means

data = as(train, "matrix")
user.means = rowMeans(data, na.rm = TRUE)
data = data - user.means
assert(dim(data) == c(M, N))

sim1 = proxy::simil(data, newdata, method = "cosine")
sim1 = (1 + sim1)/2
class(sim1)

sim1_mat <- `dim<-`(c(sim1), dim(sim1))
knn_rows = head(order(sim1[,1], decreasing = TRUE), nn)
sim_df = data.frame(
    id = rownames(sim1)[knn_rows],
    x = sim1_mat[knn_rows, 1])

# 20x3706 i.e. nn x N
knn_mat = data[knn_rows,]
assert(dim(knn_mat)[1] == nn)

# nnx1 % t(N,1) -> nn x N i.e. 20 x 3706
# Note similarity matrix might have NA in it.
s_coeff_mat = as.matrix(sim_df$x) %*% t(rep(1, N))

# Zero out the coefficient wherever the k nearest neighbors are NA.
denom_mat = s_coeff_mat * as.integer(!is.na(knn_mat))

 r = denom_mat * knn_mat
# nn x N -> vec len N
r = colSums(r, na.rm = TRUE) / colSums(denom_mat, na.rm = FALSE)
assert(length(r) == N)
mypred = r

# Clean up mypred
test_user_watched = !is.na(newdata)
assert(length(test_user_watched) == N)
mypred = mypred + newuser.means
mypred[which(is.infinite(mypred))] = NA
mypred[which(test_user_watched)] = NA
```


**Comparison to Recommenderlab**

```{r}
recommender.UBCF <- Recommender(train, method = "UBCF",
                                parameter = list(normalize = 'center', 
                                                 method = 'Cosine', 
                                                 nn = 20))

p.UBCF <- predict(recommender.UBCF, test, type="ratings")
p.UBCF <- as.numeric(as(p.UBCF, "matrix"))

sum(is.na(p.UBCF) != is.na(mypred)) ### should be zero
max(abs(p.UBCF - mypred), na.rm = TRUE)  ### should be less than 1e-06 
```






### IBCF Recommendation



**My Implementation**

```{r}
# Check this post https://campuswire.com/c/G3D46BBBA/feed/1326
# And this post https://campuswire.com/c/G3D46BBBA/feed/1334.

```

```{r}
set.seed(6723)
# center
k = 30
M = 500
N = N

newdata = as(test, "matrix")
newuser.means = mean(newdata, na.rm = TRUE)
newdata = newdata - newuser.means

data = as(train, "matrix")
user.means = rowMeans(data, na.rm = TRUE)
data = data - user.means
assert(dim(data) == c(M, N))

active_items_mask = !is.na(newdata)
active_items_idx = which(active_items_mask)

sim1 = proxy::simil(t(data), method = "cosine")
assert(dim(sim1) == c(N,N))
sim1 = (1 + sim1)/2
class(sim1)

# sim1_mat <- `dim<-`(c(sim1), dim(sim1))
sim1_mat = as.matrix(sim1)

# Remove the diag because they're all ones, which would mess up the ordering
# See office hours https://mediaspace.illinois.edu/media/t/1_2acmsvby 1:22:19
# By default empty entries in the similarity matrix should be NA not 0.
diag(sim1_mat) = NA
s_mat = matrix(NA, nrow = nrow(sim1_mat), ncol = ncol(sim1_mat))
#all_col_indices = 1:ncol(sim1_mat)
for (row in 1:nrow(sim1_mat)) {
  col_indices = head(order(sim1_mat[row,], decreasing = TRUE), k)
  col_indices = tail(order(sim1_mat[row,], na.last=FALSE, decreasing = FALSE), k)
  #inv_col_indices = setdiff(all_col_indices, col_indices)
  s_mat[row, col_indices] = sim1_mat[row, col_indices]
}
assert(dim(s_mat) == c(N,N))
tmp = !is.na(s_mat)
all(s_mat[tmp] == sim1_mat[tmp])

# Keep only k nearest neighbor rows.

# # Zero out the coefficient wherever the k nearest neighbors are NA.
# denom_mat = s_mat * as.integer(!is.na(s_mat))

# THE PROBLEM IS THIS DENOMINATOR. IT'S ALWAYS SUMMING ALL COEFFICIENTS
# WHEN WE SHOULD ONLY INCLUDE COEFFICIENTS WHEN rij != NA.
# r = (1 / sum(sim_df$x, na.rm=TRUE)) * s_coeff_mat * knn_mat 

# This is going to bring NAs into the matrix
newdata_prime = t(newdata)
r = t(t(s_mat) * as.vector(newdata_prime))
r = rowSums(r, na.rm = TRUE)
r = r / rowSums(s_mat[,active_items_idx], na.rm = TRUE)

# nn x N -> vec len N
assert(length(r) == N)
mypred = r

# Clean up mypred
mypred = mypred + newuser.means
mypred[which(is.infinite(mypred))] = NA
mypred[active_items_idx] = NA
```


**Comparison to Recommenderlab**

```{r}
recommender.IBCF <- Recommender(train, method = "IBCF",
                                parameter = list(normalize = 'center', 
                                                 method = 'Cosine', 
                                                 k = 30))

p.IBCF <- predict(recommender.IBCF, test, type="ratings")
p.IBCF <- as.numeric(as(p.IBCF, "matrix"))

## first output: should be less than 10
sum(is.na(p.IBCF) != is.na(mypred))  

## second output: should be less than 10%
mydiff = abs(p.IBCF - mypred)
sum(mydiff[!is.na(mydiff)] > 1e-6) / sum(!is.na(mydiff)) 
```


**Question: why do we encounter such a big discrepancy for IBCF, but not for
UBCF?**

IBCF results depend on the ordering of the sorted nearest neighbors when there
is a tie between two ratings. The different ordering of nearest neighbors
keeps a different column in the items matrix and later when summing over the rows
to form the prediction for a new user we get a slightly different result.






```{r eval=FALSE, echo=FALSE}
blah = predict(recommender.UBCF, test, type="topN")
# blah@itemLabels
items = as.vector(blah@items$`0`)
items
ratings = as.vector(blah@ratings$`0`)
ratings
mids = movies$MovieID %in% items
movies[mids, ]
```











